Reliable error estimation for Sobol' indices

### 2016-09-22 article/template.tex
Adding examples and plot for article
Adding all changes and new files for examples in template.tex. There is still no scrambled replicated code (cubSobol_fo_all_g)

Then, alternatively to the common Sobol' indices estimators (\ref{common.sobol.est}), we can define the following two estimators with their respective error bounds:
\begin{gather}
\begin{aligned}
\widehat{\underline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\underline{S}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_undersu}
\end{gather}
and,
\begin{gather}
\begin{aligned}
\widehat{\overline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\overline{S}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_oversu}
\end{gather}
Because numerator and denominator are known to be positive, maximizing $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) is done through maximizing the numerator $I_1$ (resp. $I_2$) and minimizing the denominator $I_3-I_4^2$. Analogously, to minimize $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) one minimizes the numerator $I_1$ (resp. $I_2$) and maximizes the denominator $I_3-I_4^2$. %As an example, Figure \ref{fig:1} below illustrate the maximization process for $\vI$ unidimensional.
%\begin{figure*}
%\centering
%\includegraphics[width=0.75\textwidth]{scheme_S.eps}
% figure caption is below the figure
%\caption{Estimator property.}
%\label{fig:1}       % Give a unique label
%\end{figure*}

@@

\section{Applications}
\label{sec:5}
\subsection{Real case model}
\label{sec:5.1}
%The Brownian motion is widely used in many applications. Thus, the following example might be interesting to understand what Sobol' indices can explain in some of these applications. For instance, if we want to estimate the expected value of the maximum of a discretized Brownian motion whose covariance matrix is $\Sigma$, our goal is to estimate
%\begin{align*}
%\mathbb{E} \left[\max(\vect{B}_t)\right] &= \int_{\mathbb{R}^d} \max(\vect{B}_t) \frac{{\rm e}^{\vect{B}_t^T\Sigma^{-1}\vect{B}_t}}{(2\pi)^{d/2}|\Sigma|^{1/2}}\,d\vect{B}_t \\
%&= \int_{[0,1]^d} \max(f_{\rm Chol}(\vx)) \,d\vx \\
%&= \int_{[0,1]^d} \max(f_{\rm PCA}(\vx)) \,d\vx.
%\end{align*}
%The last two equalities are obtained through two different subsitutions corresponding to the \textit{Cholesky} construction of the Brownian motion, and the \textit{PCA} construction. For instance, if we take 10 dimensions and $t_i=i/10$, using the Cholesky construction $S_{1} = 24\%$, $S_{2} = 16\%$, $S_{3} = 13\%$, $S_{4} = 10\%$, and $\overline{S}_{1}=24\%$, $\overline{S}_{2}=18\%$, $\overline{S}_{3}=15\%$, $\overline{S}_{4} = 13\%$. However, using the PCA construction, $S_{1} = 89\%$, $S_{2} = 2\%$, $S_{3} = 1\%$, $S_{4} = 0\%$, and $\overline{S}_{1}=94\%$, $\overline{S}_{2}=7\%$, $\overline{S}_{3}=3\%$, $\overline{S}_{4} = 2\%$. This shows that the PCA construction will be better to estimate the above expectation using quasi-Monte Carlo methods since it has a lower effective dimensionality.
%
%For those who were curious about the actual value of $\mathbb{E} \left[\max(\vect{B}_t)\right]$, it is approximately 0.5935. Nevertheless, the Brownian Motion is a continuous time process. Thus, in order to estimate the expected value on a continuous time Browninan motion, one could either use a multilevel method, or a multivariate decomposition method to estimate this infinite dimensional integral.

\subsection{Classical test functions}
\label{sec:5.2}

>>>>

Then, alternatively to the common Sobol' indices estimators (\ref{common.sobol.est}), we can define the following two estimators with their respective error bounds:
\begin{gather}
\begin{aligned}
\widehat{\underline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\underline{S}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \underline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_undersu}
\end{gather}
and,
\begin{gather}
\begin{aligned}
\widehat{\overline{S}}_u & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} + \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right) \\
\varepsilon_{\overline{S}_u} & = \frac{1}{2}\left(\min\Big{(}\max_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),1\Big{)} - \max\Big{(}\min_{\vI\in B_{\varepsilon_{\vI}}(\widehat{\vI})} \overline{S}_u(\vI),0\Big{)} \right)
\end{aligned}
\label{formula_oversu}
\end{gather}
Because numerator and denominator are known to be positive, maximizing $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) is done through maximizing the numerator $I_1$ (resp. $I_2$) and minimizing the denominator $I_3-I_4^2$. Analogously, to minimize $\underline{S}_u(\vI)$ (resp. $\overline{S}_u(\vI)$) one minimizes the numerator $I_1$ (resp. $I_2$) and maximizes the denominator $I_3-I_4^2$. As an example, Figure \ref{fig:1} below illustrate the maximization process for $\vI$ unidimensional.
\begin{figure*}
\centering
\includegraphics[width=1\textwidth]{Images/estimator_3d_small.eps}
 figure caption is below the figure
\caption{Estimator property.}
\label{fig:1}       % Give a unique label
\end{figure*}

@@

\section{Applications}
\label{sec:5}

\subsection{Classical test functions}
\label{sec:5.1}

$\varepsilon=10^{-3}$, initial $1024$ points, $fudge = @(m,d) 10*2.^-(.9*m)$, run $100$ samples to estimate the mean:

\subsubsection{With small indices estimator}
Bratley et al. example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.6530  &  0.1790  &  0.0370  &  0.0134  &  0.0016  &  0.0015 \\ \hline
 n & 521667  &    262144  &     65536  &  17039  &   1055  &      1096 \\ \hline
 \text{error } 10^{-3} & 0.0211  &  0.0520  &  0.1160  &  0.1550  &  0.3220  &  0.3210 \\ \hline
 \widehat{\overline{S}}_u & 0.7400  &  0.2660  &  0.0764  &  0.0343  &  0.0060  &  0.0061 \\  \hline
 n & 262144  &    131072  &   65536  &   40960  &  4936  &      6031 \\ \hline
 \text{error } 10^{-3} & 0.0578  &  0.0374  &  0.0761  &  0.1820  &  0.5530  &  0.2740 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 100  & 100  &  99  & 100 \\
     100  & 100  & 100  &  99  &  84  &  90
\end{array}
\]

Sobol g-function example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.5870  &  0.2610  &  0.0365 &   0.0059  &  0.0001  &  0.0001 \\ \hline
 n & 131072   &  130417  &  17695  &  1987  &  1024  &   1024 \\ \hline
 \text{error } 10^{-3} & 0.0235  &  0.0228  &  0.2140 & 0.1560  &  0.0026  &  0.0023 \\ \hline
 \widehat{\overline{S}}_u & 0.6900  &  0.3560  &  0.0563 &   0.0092  &  0.0001  &  0.0001 \\  \hline
 n & 110756  &  65536  &  16056  &  2550  &  1024  &  1024 \\ \hline
 \text{error } 10^{-3} & 0.0366 &  0.0413  &  0.0856 & 0.0838  &  0.0022  &  0.0028 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 93  & 100  &  100  & 100 \\
     98  &  99  &  98 &  100  & 100 &  100
\end{array}
\]

\subsubsection{Without small indices estimator}
Bratley et al. example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.6530 & 0.1790 & 0.0370 & 0.0133 & 0.0014 & 0.0015 \\ \hline
 n & 524288 & 262144  &  222822 & 131072 & 57190 & 64901 \\ \hline
 \text{error } 10^{-3} & 0.0275 & 0.0592 & 0.0536 & 0.0667 &  0.2030 & 0.1180 \\ \hline
 \widehat{\overline{S}}_u & 0.7400 & 0.2660 & 0.0764 & 0.0343 & 0.0060 & 0.0062 \\  \hline
 n & 262144 & 131072 &  65536 &  39322 & 4710 & 6339 \\ \hline
 \text{error } 10^{-3} & 0.0781 & 0.0371 & 0.1090 & 0.2150 &  0.4850 & 0.1740 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 100  & 100  &  97  & 99 \\
     100  & 100  & 100  &  99  &  88  &  98
\end{array}
\]

Sobol g-function example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.5870 & 0.2610 & 0.0367 &  0.0059 & 0.0004 & 0.0004 \\ \hline
 n & 131072 & 131072 & 63898 & 19825 & 1024 & 1024 \\ \hline
 \text{error } 10^{-3} & 0.0280 & 0.0262 & 0.0476 & 0.0736 & 0.3390 & 0.3440 \\ \hline
 \widehat{\overline{S}}_u & 0.6900 & 0.3560 & 0.0563 & 0.0092 & 0.0001 &  0.0001 \\  \hline
 n & 105513 & 65536 & 16056 & 2703 & 1024 & 1024 \\ \hline
 \text{error } 10^{-3} & 0.0507 & 0.0442 & 0.1270 & 0.0703 & 0.0022 & 0.0027 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 99  & 100  &  100  & 100 \\
     98  &  99  &  97 &  100  & 100 &  100
\end{array}
\]

\subsubsection{Replicated method}
Bratley et al. example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.6530 & 0.1790 & 0.0370 & 0.0133 & 0.0014 & 0.0015 \\ \hline
 n & 524288 & 262144  &  222822 & 131072 & 57190 & 64901 \\ \hline
 \text{error } 10^{-3} & 0.0275 & 0.0592 & 0.0536 & 0.0667 &  0.2030 & 0.1180 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 100  & 100  &  97  & 99
\end{array}
\]

Sobol g-function example:

\[
\begin{array}{lcccccc}
   & u=\{1\} & u=\{2\} & u=\{3\} & u=\{4\} & u=\{5\} & u=\{6\} \\ \hline \hline
 \widehat{\underline{S}}_u & 0.5870 & 0.2610 & 0.0367 &  0.0059 & 0.0004 & 0.0004 \\ \hline
 n & 131072 & 131072 & 63898 & 19825 & 1024 & 1024 \\ \hline
 \text{error } 10^{-3} & 0.0280 & 0.0262 & 0.0476 & 0.0736 & 0.3390 & 0.3440 \\ \hline
\end{array}
\]
Reliability on the results,
\[
\begin{array}{ cccccc }
     100  & 100  & 99  & 100  &  100  & 100
\end{array}
\]

\subsection{Real case model}
\label{sec:5.2}
%The Brownian motion is widely used in many applications. Thus, the following example might be interesting to understand what Sobol' indices can explain in some of these applications. For instance, if we want to estimate the expected value of the maximum of a discretized Brownian motion whose covariance matrix is $\Sigma$, our goal is to estimate
%\begin{align*}
%\mathbb{E} \left[\max(\vect{B}_t)\right] &= \int_{\mathbb{R}^d} \max(\vect{B}_t) \frac{{\rm e}^{\vect{B}_t^T\Sigma^{-1}\vect{B}_t}}{(2\pi)^{d/2}|\Sigma|^{1/2}}\,d\vect{B}_t \\
%&= \int_{[0,1]^d} \max(f_{\rm Chol}(\vx)) \,d\vx \\
%&= \int_{[0,1]^d} \max(f_{\rm PCA}(\vx)) \,d\vx.
%\end{align*}
%The last two equalities are obtained through two different subsitutions corresponding to the \textit{Cholesky} construction of the Brownian motion, and the \textit{PCA} construction. For instance, if we take 10 dimensions and $t_i=i/10$, using the Cholesky construction $S_{1} = 24\%$, $S_{2} = 16\%$, $S_{3} = 13\%$, $S_{4} = 10\%$, and $\overline{S}_{1}=24\%$, $\overline{S}_{2}=18\%$, $\overline{S}_{3}=15\%$, $\overline{S}_{4} = 13\%$. However, using the PCA construction, $S_{1} = 89\%$, $S_{2} = 2\%$, $S_{3} = 1\%$, $S_{4} = 0\%$, and $\overline{S}_{1}=94\%$, $\overline{S}_{2}=7\%$, $\overline{S}_{3}=3\%$, $\overline{S}_{4} = 2\%$. This shows that the PCA construction will be better to estimate the above expectation using quasi-Monte Carlo methods since it has a lower effective dimensionality.
%
%For those who were curious about the actual value of $\mathbb{E} \left[\max(\vect{B}_t)\right]$, it is approximately 0.5935. Nevertheless, the Brownian Motion is a continuous time process. Thus, in order to estimate the expected value on a continuous time Browninan motion, one could either use a multilevel method, or a multivariate decomposition method to estimate this infinite dimensional integral.
